A Decision Procedure for (Co)datatypes in SMT Solvers

### 2015-02-19 tex/body.tex
fonts

@@

\begin{rep}\paragraph{Related Work.} Despite\end{rep}%
\begin{conf}Despite\end{conf} its age,
Barrett et al.\ \cite{barrett-et-al-2007} still provides a good account of
related work. Since then, datatypes have been added to the SMT solver
Z3 \cite{de-moura-bjoerner-2008} \begin{rep}in unpublished work by Leonardo de Moura \end{rep}and to a SPASS-like %prototypical
superposition prover \begin{rep}by Daniel Wand\end{rep} \cite{wand-2014}.
Closely related are the automatic structural
induction in both kinds of provers \cite{kersani-peltier-2013,cruanes-201x,wand-weidenbach-201x,reynolds-kuncak-2015},
the (co)datatype and (co)induction support in Dafny \cite{leino-moskal-2014},
and the (semi-)decision procedures for datatypes
implemented in Leon \cite{suter-et-al-2011} and RADA \cite{pham-et-al-2013}.
%
Datatypes are supported by the higher-order model finder Refute
\cite{weber-2008}. Its successor, Nitpick \cite{blanchette-nipkow-2010}, can
also generate models involving cyclic codatatype values. The $\mu$-notation is
inspired by the $\mu$-calculus
%\begin{conf}\cite{kozen-1983}.\end{conf}%
%\begin{rep}\cite[etc.]{kozen-1983,endrullis-et-al-2011}.\end{rep}%
\cite[etc.]{kozen-1983,endrullis-et-al-2011}.

@@

\paragraph{Conventions.}
The setting is a monomorphic (or many-sorted) first-order logic.
A signature $\Sig = (\Types, \Funcs)$ consists of a set of types $\Types$ and a
set of function symbols $\Funcs$. Types are atomic sorts and interpreted as
nonempty domains. The set~$\Types$ must contain a
\begin{rep}distinguished \end{rep}type \ty{bool} interpreted as the set of truth
values. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
\begin{rep}
The metavariables $\delta,\:\varepsilon$ range over (co)datatypes,
whereas $\tau,\:\upsilon$ range over arbitrary types.
When applied to terms, the symbol $=$ denotes syntactic equality.

@@

%\paragraph{Specifications.}
In an SMT problem, the signature is typically given by specifying first the
uninterpreted ordinary types in any order, then the (co)datatypes with their constructors
and selectors in groups of mutually (co)recursive groups of (co)datatypes, and
finally any other function symbols.
%
Each (co)datatype specification consists of $l$~mutually recursive types that are
either all datatypes or all codatatypes. Polymorphic types, nested
(co)recursion, and datatype--codatatype mixtures fall outside this fragment%
\dotReportFootnote{%
In principle, rank-1 (top-level) polymorphism \cite{blanchette-paskevich-2013}
should not raise any special difficulties. Nesting datatypes inside datatypes,
and likewise for codatatypes, can be reduced to the mutual case
\cite{gunter-1993-not}. So the only genuinely interesting cases missing are
mixed nested (co)recursion as well as
(co)recursion through a non-(co)datatype (both of which make sense
\cite{blanchette-et-al-2014-codata}).}
We allow ourselves some notational parameterization
% at the metalevel
through subscripts%
\begin{conf}
(e.g., $\ty{stream}_{\,\tau}$)%
\end{conf}%
\begin{rep}---for example, $\ty{stream}_{\,\tau}$ denotes a
family of ground types including
$\ty{stream}_{\,\ty{int}}$, $\ty{stream}_{\,\ty{bool}}$,
and \smash{$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$}\end{rep}.

@@

%\paragraph{Characterization.}
One way to define datatypes semantically is as the initial model of the
selector--constructor equations \cite{barrett-et-al-2007}.
\begin{rep}
A drawback of this approach is that it does
not naturally account for selectors applied to wrong constructors. Barrett et
al.\ address this by parameterizing the construction by default values, but
this gives rise to spurious equalities between unrelated terms---e.g.,
$\const{s}_1^1(\const{C}_2) \teq \const{s}_1^1(\const{C}_3)$. This flaw
could be corrected, but the added complexity seems to suggest that selectors
are better characterized axiomatically.

@@

%\paragraph{Special Cases.}
{\looseness=-1
Some codatatypes are so degenerate as to have infinite values
even though they are finite. A simple example is
\vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, whose
unique value is $\MU a.\; \const{A}(a)$. Other specimens are
$\ty{stream}_{\,\ty{unit}\,}$
and
\begin{conf}%
\vthinspace$\keyw{codatatype}~\ty{b} = \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit})
~\keyw{and}~ \ty{c} = \const{C}(\ty{a},\: \ty{unit},\: \ty{b},\: \ty{c})$,
\end{conf}%
\begin{rep}%
\[\begin{aligned}[t]
      \keyw{codatatype}\;\, \ty{b} & \,=\, \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit}) \\[-1\jot]
      \keyw{and}\;\, \ty{c} & \,=\, \const{C}(\ty{a},\: \ty{unit},\: \ty{b},\: \ty{c})
\end{aligned}
\]
\end{rep}%
where \ty{unit} is a datatype with the single constructor $\const{Unity} :
\ty{unit}$. We call such types
\emph{corecursive singletons}. For the decision procedure, it will be
crucial to detect these. %, even if they rarely arise in practice.
A type may also be a corecursive singleton only in some models. If the example
above is altered to make \ty{unit} an ordinary type, \ty{b} and \ty{c} will be
singletons precisely when \ty{unit} is interpreted as a singleton.
Fortunately, \begin{rep}given cardinalities for the ordinary types, \end{rep}%
it is easy to characterize this degenerate case.

@@

\paragraph{Phase 1: Computing the Bidirectional Closure \rm(\figurename~\ref{fig:cc-rules}).}
In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the (upward) congruence closure,
whereas the \rn{Inject} and \rn{Clash} rules together compute (downward) unification.
For unification, additional equalities are inferred based on the injectivity of constructors by \rn{Inject},
and failures to unify equated terms are recognized by \rn{Clash}.
The \rn{Conflict} rule recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.

@@

\paragraph{Phase 2: Applying Acyclicity and Uniqueness \rm(\figurename~\ref{fig:ab-rules}).}
%For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
The premises of the rules that make up the second phase refer to a mapping $\Val$
that associates a $\mu$-term with each equivalence class as its representative.
%Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
%(Section~\ref{sec:introduction}).
% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
The $\mu$-term $\Val[t\typ{\tau}]$ describes a class of $\tau$ values
that $t$ and other members of $t$'s equivalence class can take in models of $\Ec$.
%either in the case when $\tau$ is a codatatype or datatype type,
When $\tau$ is a datatype, % \in \Data$,
a cyclic $\mu$-term describes an infeasible class of values.
%$\tau \not\in \Codata$.

@@

\paragraph{Phase 3: Branching \rm(\figurename~\ref{fig:split-rule}).}
If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
$t$'s equivalence class must contain a
$\delta$ constructor term.
This is enforced in the third phase by the \rn{Split} rule.
Another rule, \rn{Single}, focuses on the degenerate case where two
terms %from $\tEc$
have a corecursive singleton type
%(Section~\ref{sec:the-theory-of-co-datatypes})
and must therefore be equal. Both
\rn{Split}'s finiteness assumption %on the datatype
and \rn{Single}'s singleton
constraint %on the codatatype
can be evaluated statically
based on a recursive computation of the cardinalities of the
constructors' argument types.
%This step relies on the assumption that ordinary types are infinite.

\paragraph{Correctness.}
Correctness means that if there exists a closed derivation tree with root node
$\Ec$, then $\Ec$ is $\thD$-unsatisfiable; and if there exists
a derivation tree with root node $\Ec$ that contains a saturated node, then
$\Ec$ is $\thD$-satisfiable.

@@

%%% @ANDY: I'm trying without \paragraph's, to save space but also to make the
%%% section look more like a whole than a random list of stuff. Revert if you
%%% disagree.

%\paragraph{Integration into {\relax\XXXL($T\negvthinspace$)}.}
{\looseness=-1
The calculus is implemented as a \relax{theory solver} of CVC4, that is, a
specialized solver for determining the satisfiability of conjunctions of
constraints for its theory.
Given a theory $T = T_1 \mathrel\cup \cdots \mathrel\cup T_n$ and a set of
input clauses $F$ in conjunctive normal form, the \XXXL($T$) procedure
incrementally builds partial assignments from the atomic formulas of $F$ to
truth values such that no clause in $F$ is falsified. We can regard such a
partial assignment as a set $M$ of true literals.
% where $a \in M$ for all atoms $a$ that
%it assigned the value $\top$ and $\neg\; a \in M$ for all atoms that it assigns
%the value $\bot$.
%at a high level,
By a variant
\cite{jovanovic2011sharing} of the Nelson--Oppen method,
each $T_i$-solver %for $i = 1, \ldots, l$  n?
takes as input the union $M_i$
of (1)~the purified form of $T_i$-constraints occurring in $M$, where terms
containing symbols not belonging to $T_i$ are replaced by fresh variables\begin{rep} of the
same type\end{rep}; (2)~additional (dis)equalities between variables of
types not belonging to $T_i$.
Each $T_i$-solver
either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
$\neg\; C$ is added to $F\!$, adds a clause to $F\!$, or does nothing.
When $M$ is a complete assignment for $F\!$, a theory solver can choose to do
nothing only if $M_i$ is indeed $T_i$-satisfiable.

@@

%\paragraph{Extension to Quantified Formulas.}
While the decision procedure is restricted to ground constraints,
in practice we often want to solve problems that feature universal axioms and existential conjectures,
where checking satisfiability requires reasoning about such ground constraints in the presence of universally quantified formulas.
Many SMT solvers, including CVC4, can reason about quantified formulas using incomplete instantiation-based methods~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
These methods extend naturally to %quantified
formulas involving datatypes and codatatypes.

@@

%\paragraph{Optimizations.}
The implementation of the decision procedure uses several optimizations
following the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
%and are not reflected in the presentation of the calculus.
\begin{rep}We briefly mention the main ones. \end{rep}%
Discriminators are part of the signature and not abbreviations.
This requires extending the decision procedure with several
rules \cite{barrett-et-al-2007},
which apply uniformly to datatypes and codatatypes.
This approach often leads to better performance because it introduces terms
less eagerly to $\tEc$.
Selectors are collapsed eagerly:
If $\const s^k_{\negvthinspace j}( t ) \in \tEc$ and $t =
\const{C}_{\negvthinspace j}( u_1, \ldots, u_n )$, the solver directly adds
$\smash{\const s^k_{\negvthinspace j}( t ) \teq u_k}$ to $\Ec$, whereas the presented calculus
would apply \rn{Split} and \rn{Inject} before adding this equality.

@@

\def\ackname{Acknowledgment}
\paragraph{\ackname.}
We owe a great debt to the development team of CVC4, 
including Clark Barrett and Cesare Tinelli,
and in particular Morgan Deters,
who jointly with the first author developed the initial version 
of the theory solver for datatypes in CVC4.
%
Our present and former bosses, Viktor Kuncak, Stephan Merz, Tobias Nipkow,
Cesare Tinelli, and Christoph Weidenbach, have either encouraged the research on
codatatypes or at least benevolently tolerated it, both of which we are thankful
for.
%
Peter Gammie and Andreas Lochbihler shared their private
theories with us so we could include them in the benchmarks.
Andrei Popescu helped clarify our thoughts regarding the axiomatization of
codatatypes. Dmitriy Traytel took part in discussions about degenerate
codatatypes.
%
Andreas Lochbihler and Mark Summerfield suggested several textual improvements.
The second author's research was partially supported by the Deutsche
Forschungs\-gemein\-schaft (DFG) project
\relax{Hardening the Hammer} (grant Ni\,491\slash 14-1).

>>>>

%\newcommand\ourparagraph[1]{\paragraph{\upshape\textbf{#1}}}
%\newcommand\beforefig{}

\newcommand\ourparagraph[1]{\paragraph{{#1}}}
\newcommand\beforefig{\rm}

@@

\begin{rep}\ourparagraph{Related Work.} Despite\end{rep}%
\begin{conf}Despite\end{conf} its age,
Barrett et al.\ \cite{barrett-et-al-2007} still provides a good account of
related work. Since then, datatypes have been added to the SMT solver
Z3 \cite{de-moura-bjoerner-2008} \begin{rep}in unpublished work by Leonardo de Moura \end{rep}and to a SPASS-like %prototypical
superposition prover \begin{rep}by Daniel Wand\end{rep} \cite{wand-2014}.
Closely related are the automatic structural
induction in both kinds of provers \cite{kersani-peltier-2013,cruanes-201x,wand-weidenbach-201x,reynolds-kuncak-2015},
the (co)datatype and (co)induction support in Dafny \cite{leino-moskal-2014},
and the (semi-)decision procedures for datatypes
implemented in Leon \cite{suter-et-al-2011} and RADA \cite{pham-et-al-2013}.
%
Datatypes are supported by the higher-order model finder Refute
\cite{weber-2008}. Its successor, Nitpick \cite{blanchette-nipkow-2010}, can
also generate models involving cyclic codatatype values. The $\mu$-notation is
inspired by the $\mu$-calculus
%\begin{conf}\cite{kozen-1983}.\end{conf}%
%\begin{rep}\cite[etc.]{kozen-1983,endrullis-et-al-2011}.\end{rep}%
\cite[etc.]{kozen-1983,endrullis-et-al-2011}.

@@

\ourparagraph{Conventions.}
The setting is a monomorphic (or many-sorted) first-order logic.
A signature $\Sig = (\Types, \Funcs)$ consists of a set of types $\Types$ and a
set of function symbols $\Funcs$. Types are atomic sorts and interpreted as
nonempty domains. The set~$\Types$ must contain a
\begin{rep}distinguished \end{rep}type \ty{bool} interpreted as the set of truth
values. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
\begin{rep}
The metavariables $\delta,\:\varepsilon$ range over (co)datatypes,
whereas $\tau,\:\upsilon$ range over arbitrary types.
When applied to terms, the symbol $=$ denotes syntactic equality.

@@

%\ourparagraph{Specifications.}
In an SMT problem, the signature is typically given by specifying first the
uninterpreted ordinary types in any order, then the (co)datatypes with their constructors
and selectors in groups of mutually (co)recursive groups of (co)datatypes, and
finally any other function symbols.
%
Each (co)datatype specification consists of $l$~mutually recursive types that are
either all datatypes or all codatatypes. Polymorphic types, nested
(co)recursion, and datatype--codatatype mixtures fall outside this fragment%
\dotReportFootnote{%
In principle, rank-1 (top-level) polymorphism \cite{blanchette-paskevich-2013}
should not raise any special difficulties. Nesting datatypes inside datatypes,
and likewise for codatatypes, can be reduced to the mutual case
\cite{gunter-1993-not}. So the only genuinely interesting cases missing are
mixed nested (co)recursion as well as
(co)recursion through a non-(co)datatype (both of which make sense
\cite{blanchette-et-al-2014-codata}).}
We allow ourselves some notational parameterization
% at the metalevel
through subscripts%
\begin{conf}
(e.g., $\ty{stream}_{\,\tau}$)%
\end{conf}%
\begin{rep}---for example, $\ty{stream}_{\,\tau}$ denotes a
family of ground types including
$\ty{stream}_{\,\ty{int}}$, $\ty{stream}_{\,\ty{bool}}$,
and \smash{$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$}\end{rep}.

@@

%\ourparagraph{Characterization.}
One way to define datatypes semantically is as the initial model of the
selector--constructor equations \cite{barrett-et-al-2007}.
\begin{rep}
A drawback of this approach is that it does
not naturally account for selectors applied to wrong constructors. Barrett et
al.\ address this by parameterizing the construction by default values, but
this gives rise to spurious equalities between unrelated terms---e.g.,
$\const{s}_1^1(\const{C}_2) \teq \const{s}_1^1(\const{C}_3)$. This flaw
could be corrected, but the added complexity seems to suggest that selectors
are better characterized axiomatically.

@@

%\ourparagraph{Special Cases.}
{\looseness=-1
Some codatatypes are so degenerate as to have infinite values
even though they are finite. A simple example is
\vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, whose
unique value is $\MU a.\; \const{A}(a)$. Other specimens are
$\ty{stream}_{\,\ty{unit}\,}$
and
\begin{conf}%
\vthinspace$\keyw{codatatype}~\ty{b} = \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit})
~\keyw{and}~ \ty{c} = \const{C}(\ty{a},\: \ty{unit},\: \ty{b},\: \ty{c})$,
\end{conf}%
\begin{rep}%
\[\begin{aligned}[t]
      \keyw{codatatype}\;\, \ty{b} & \,=\, \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit}) \\[-1\jot]
      \keyw{and}\;\, \ty{c} & \,=\, \const{C}(\ty{a},\: \ty{unit},\: \ty{b},\: \ty{c})
\end{aligned}
\]
\end{rep}%
where \ty{unit} is a datatype with the single constructor $\const{Unity} :
\ty{unit}$. We call such types
\emph{corecursive singletons}. For the decision procedure, it will be
crucial to detect these. %, even if they rarely arise in practice.
A type may also be a corecursive singleton only in some models. If the example
above is altered to make \ty{unit} an ordinary type, \ty{b} and \ty{c} will be
singletons precisely when \ty{unit} is interpreted as a singleton.
Fortunately, \begin{rep}given cardinalities for the ordinary types, \end{rep}%
it is easy to characterize this degenerate case.

@@

\ourparagraph{Phase 1: Computing the Bidirectional Closure \beforefig(\figurename~\ref{fig:cc-rules}).}
In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the (upward) congruence closure,
whereas the \rn{Inject} and \rn{Clash} rules together compute (downward) unification.
For unification, additional equalities are inferred based on the injectivity of constructors by \rn{Inject},
and failures to unify equated terms are recognized by \rn{Clash}.
The \rn{Conflict} rule recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.

@@

\ourparagraph{Phase 2: Applying Acyclicity and Uniqueness \beforefig(\figurename~\ref{fig:ab-rules}).}
%For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
The premises of the rules that make up the second phase refer to a mapping $\Val$
that associates a $\mu$-term with each equivalence class as its representative.
%Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
%(Section~\ref{sec:introduction}).
% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
The $\mu$-term $\Val[t\typ{\tau}]$ describes a class of $\tau$ values
that $t$ and other members of $t$'s equivalence class can take in models of $\Ec$.
%either in the case when $\tau$ is a codatatype or datatype type,
When $\tau$ is a datatype, % \in \Data$,
a cyclic $\mu$-term describes an infeasible class of values.
%$\tau \not\in \Codata$.

@@

\ourparagraph{Phase 3: Branching \beforefig(\figurename~\ref{fig:split-rule}).}
If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
$t$'s equivalence class must contain a
$\delta$ constructor term.
This is enforced in the third phase by the \rn{Split} rule.
Another rule, \rn{Single}, focuses on the degenerate case where two
terms %from $\tEc$
have a corecursive singleton type
%(Section~\ref{sec:the-theory-of-co-datatypes})
and must therefore be equal. Both
\rn{Split}'s finiteness assumption %on the datatype
and \rn{Single}'s singleton
constraint %on the codatatype
can be evaluated statically
based on a recursive computation of the cardinalities of the
constructors' argument types.
%This step relies on the assumption that ordinary types are infinite.

\ourparagraph{Correctness.}
Correctness means that if there exists a closed derivation tree with root node
$\Ec$, then $\Ec$ is $\thD$-unsatisfiable; and if there exists
a derivation tree with root node $\Ec$ that contains a saturated node, then
$\Ec$ is $\thD$-satisfiable.

@@

%%% @ANDY: I'm trying without \ourparagraph's, to save space but also to make the
%%% section look more like a whole than a random list of stuff. Revert if you
%%% disagree.

%\ourparagraph{Integration into {\relax\XXXL($T\negvthinspace$)}.}
{\looseness=-1
The calculus is implemented as a \relax{theory solver} of CVC4, that is, a
specialized solver for determining the satisfiability of conjunctions of
constraints for its theory.
Given a theory $T = T_1 \mathrel\cup \cdots \mathrel\cup T_n$ and a set of
input clauses $F$ in conjunctive normal form, the \XXXL($T$) procedure
incrementally builds partial assignments from the atomic formulas of $F$ to
truth values such that no clause in $F$ is falsified. We can regard such a
partial assignment as a set $M$ of true literals.
% where $a \in M$ for all atoms $a$ that
%it assigned the value $\top$ and $\neg\; a \in M$ for all atoms that it assigns
%the value $\bot$.
%at a high level,
By a variant
\cite{jovanovic2011sharing} of the Nelson--Oppen method,
each $T_i$-solver %for $i = 1, \ldots, l$  n?
takes as input the union $M_i$
of (1)~the purified form of $T_i$-constraints occurring in $M$, where terms
containing symbols not belonging to $T_i$ are replaced by fresh variables\begin{rep} of the
same type\end{rep}; (2)~additional (dis)equalities between variables of
types not belonging to $T_i$.
Each $T_i$-solver
either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
$\neg\; C$ is added to $F\!$, adds a clause to $F\!$, or does nothing.
When $M$ is a complete assignment for $F\!$, a theory solver can choose to do
nothing only if $M_i$ is indeed $T_i$-satisfiable.

@@

%\ourparagraph{Extension to Quantified Formulas.}
While the decision procedure is restricted to ground constraints,
in practice we often want to solve problems that feature universal axioms and existential conjectures,
where checking satisfiability requires reasoning about such ground constraints in the presence of universally quantified formulas.
Many SMT solvers, including CVC4, can reason about quantified formulas using incomplete instantiation-based methods~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
These methods extend naturally to %quantified
formulas involving datatypes and codatatypes.

@@

%\ourparagraph{Optimizations.}
The implementation of the decision procedure uses several optimizations
following the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
%and are not reflected in the presentation of the calculus.
\begin{rep}We briefly mention the main ones. \end{rep}%
Discriminators are part of the signature and not abbreviations.
This requires extending the decision procedure with several
rules \cite{barrett-et-al-2007},
which apply uniformly to datatypes and codatatypes.
This approach often leads to better performance because it introduces terms
less eagerly to $\tEc$.
Selectors are collapsed eagerly:
If $\const s^k_{\negvthinspace j}( t ) \in \tEc$ and $t =
\const{C}_{\negvthinspace j}( u_1, \ldots, u_n )$, the solver directly adds
$\smash{\const s^k_{\negvthinspace j}( t ) \teq u_k}$ to $\Ec$, whereas the presented calculus
would apply \rn{Split} and \rn{Inject} before adding this equality.

@@

\def\ackname{Acknowledgment}
\ourparagraph{\ackname.}
We owe a great debt to the development team of CVC4, 
including Clark Barrett and Cesare Tinelli,
and in particular Morgan Deters,
who jointly with the first author developed the initial version 
of the theory solver for datatypes in CVC4.
%
Our present and former bosses, Viktor Kuncak, Stephan Merz, Tobias Nipkow,
Cesare Tinelli, and Christoph Weidenbach, have either encouraged the research on
codatatypes or at least benevolently tolerated it, both of which we are thankful
for.
%
Peter Gammie and Andreas Lochbihler shared their private
theories with us so we could include them in the benchmarks.
Andrei Popescu helped clarify our thoughts regarding the axiomatization of
codatatypes. Dmitriy Traytel took part in discussions about degenerate
codatatypes.
%
Andreas Lochbihler and Mark Summerfield suggested several textual improvements.
The second author's research was partially supported by the Deutsche
Forschungs\-gemein\-schaft (DFG) project
\relax{Hardening the Hammer} (grant Ni\,491\slash 14-1).
